---
title: "R Packages"
subtitle: "Or, The Best Way to Distribute Anything Ever"  
author: "Nick Solomon | New York City Council"
date: 'October 19, 2018'
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: xaringan-themer.css
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```

```{r xaringan-themer, include=FALSE}
library(xaringanthemer)

mono_light(
  base_color = "#2f56a6",
  white_color = "#FFFFFF",
  black_color = "#000000",
  text_font_size = "28px",
  text_font_google = google_font("Lato", "400", "400i", "700", "700i"),
  header_font_google = google_font("Roboto Slab"),
  header_color = darken_color("#2f56a6", .3),
  code_inline_color = darken_color("#2f56a6", .4),
  code_highlight_color = "rgba(88, 119, 183, .5)",
  text_slide_number_font_size = ".7em"
)
```

class: center, middle, inverse

# What is an R package?

---

## What is an R package?
--

- Code
--

- Documentation
--

- Data
--

- Vignettes

---

## What is an R package?
--

- `R/`
--

- `man/`
--

- `data/`
--

- `vignettes/`
--

- `DESCRIPTION`
- `NAMESPACE`
- `LICENSE`
- `tests/`
---

## The package paradigm
--

If your project is a package then:
--

- Everything is versioned
--

- Dependencies are explicitly stated
--

  - In a machine readable way
--

- The whole project can easily be shared
--

- You can be almost certain the project is reproducible

---

## When to create an R package?
--

- When you write code that:
--

  - Will be shared
--

  - Will be run more than once
--

  - Will be run by someone else
--

  - Should be well documented
--

- When you have data that:
--

  - Will be shared
--

  - Has been cleaned
--

- When you write reports that:
--

  - Need to be reproducible
--

  - Will be shared

---

## Disadvantages of creating a package
--

- Administrative overhead
--

- A new workflow
--

- Forces best practices `r emo::ji("wink")`

---

class: center, middle, inverse

# Code

---

class: center, middle, inverse

# Documentation

---

class: center, middle, inverse

# Data

---

class: center, middle, inverse

# Vignettes

---

class: center, middle, inverse

# Resources
